---
title: "PROJECT TITLE"
author: "TEAM NAME"
date: "TODAY'S DATE"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(broom)
library(rvest)
```


```{r load-data, echo=FALSE}
players_combined <- read.csv("data/players_combined.csv")
glimpse(players_combined)
```

```{r summary-statistics, echo=FALSE}
players_combined %>%
  count(position)
```

We can see that the most common position among the top 50 players is Centre-Forward. Most players are attackers or midfielders instead of defenders.

```{r age}
players_combined <- players_combined %>%
  mutate(age_range = case_when(
    age>30  ~ "above 30",
    26<=age & age<=30     ~ "26-30",
    21<=age & age<=25     ~ "21-25",
    age<=20  ~ "under 20"
  ))
players_combined %>%
  count(age_range)
```

We have tried to classify the players by age. Most players age between 26 and 30. There are 3 players above 30 years old and there's also one player who is younger than 20.

```{r matches, echo=FALSE}
players_combined %>%
  summarise(mean_matches = mean(matches))
players_combined %>%
  summarise(sd_matches = sd(matches))
```

The average matches played during season 2018-2019 by these top 50 players is 17.34 and the standard deviation is 3.51.

```{r goals, echo=FALSE}
players_combined %>%
  summarise(mean_goals = mean(goals))
players_combined %>%
  summarise(sd_goals = sd(goals))
```

The average goals scored during season 2018-2019 by these top 50 players is 5.22 and the standard deviation is 4.51.

```{r assists, echo=FALSE}
players_combined %>%
  summarise(mean_assists = mean(assists))
players_combined %>%
  summarise(sd_assists = sd(assists))
```

The average assists made during season 2018-2019 by these top 50 players is 3.04 and the standard deviation is 2.73.

```{r yellow_cards, echo=FALSE}
players_combined %>%
  count(yellow_cards)
```

We can see that until now, 38 among the 50 players have got yellow cards during season 2018-2019, and the maximum yellow card number is 5.

```{r red_card, echo=FALSE}
players_combined %>%
  filter(red_cards!=0)
```

In season 2018-2019, there are 3 players who have already got red card in a match and being expelled: Kylian Mbappé, Cristiano Ronaldo, and Marcus Rashford.

```{r market_value}
players_combined %>%
  summarise(max_value = max(market_value))
players_combined %>%
  summarise(min_value = min(market_value))
players_combined %>%
  filter(market_value>=100) %>%
  summarise(value_above_100m = n())
```

In the dataset, the maximum player value is 180 million euro and the minimum player value is 65 million euro. There are 13 players in the dataset with a market value higher than 100 million euro.


Your project goes here! Before you submit, make sure your chunks are turned off with `echo = FALSE`. 

You can add sections as you see fit. Make sure you have a section called Introduction at the beginning and a section called Conclusion at the end. The rest is up to you!
