---
title: "Building the Perfect Football Player"
author: "team-devils"
date: "9/19/2018"
output: github_document
---

```{r load_packages, include = FALSE}
library(tidyverse)
library(broom)
library(rvest)
```


## Section 1. Introduction

We wish to explore how performance, physical attributes, and general characteristics of soccer players impact their market value during a season. Through visualizations and modeling, we hope to explore which characteristics are more and less determinant of strong performance and market value for a player, and whether patterns exist within goals, assists, age ranges, positions, and more. We plan to first minimally use visualizations to discover patterns, and then modeling to develop our central questions of which characteristics of players impact performance and value. 

Our data is a collection of the top 50 valuable players in the 2018-2019 European club season. The data comes from a professional German soccer statistics website titled "Transfermarkt", which is a website dedicated to tracking player's market values and performances. Each observation is a player, and includes the variables name, position, number of matches, number of goals scored, number of own goals, number of assists, number of yellow cards, number of red cards, number of substitutions on, number of substitutions off, and market value. The data was obtained from transfermarkt.com through web scraping tools learned from the course, and the web_scrape.R file can be found in this R project. We currently have 12 columns and 50 rows scraped, but may later during our analysis decide to obtain a few more columns and several more rows, depending on where our analysis leads us.

Source of data: Transfermarkt.com is a leading media in reporting soccer transfer news and they have connections with all the major leagues and clubs across Europe, South America, and Asia. The player statistics are generated after each match and analyzed by professional scouts and soccer analysts.

## Section 2. Data analysis plan

## Variables

Here we use 11 variables as predictors (position, age, matches, goals, own_goals, assists, yellow_cards, red_cards, substituted_on, substituted_off). These variables are used to predict the final outcome variable ---- market_value. The meanings of all variables are explained in the codebook in README.

## Statistical Methods

We will first test to see if each variable makes a significant difference in the market value of a player using hypotheses testing. For the variables that show significant correlation, we will also calculate a confidence interval using bootstrapping. Then, we will make a model for `market_value`, using all of the explanatory variables and the possible interactions between them (e.g. the interaction between `position` and `goals` because common knowledge tells us that there will be a correlation between the two). After we make the model, we will perform a backwards selection to remove the unnecessary variables. We will test the final model on partitions of our data to evaluate it, and in the end attempt to come up with a generalization about the conditions that will produce the highest market value.

```{r load-data, echo=FALSE}
players_combined <- read.csv("players_combined.csv")
```

### Summary Statistics

```{r summary-statistics, echo=FALSE}
players_combined %>%
  count(position)
```

We can see that the most common position among the top 50 players is Centre-Forward. Most players are attackers or midfielders instead of defenders.

```{r age, echo=FALSE}
players_combined <- players_combined %>%
  mutate(age_range = case_when(
    age>30  ~ "above 30",
    26<=age & age<=30     ~ "26-30",
    21<=age & age<=25     ~ "21-25",
    age<=20  ~ "under 20"
  ))
players_combined %>%
  count(age_range)
```

We have tried to classify the players by age. Most players age between 26 and 30. There are 3 players above 30 years old and there's also one player who is younger than 20.

```{r matches, echo=FALSE}
players_combined %>%
  summarise(mean_matches = mean(matches), sd_matches = sd(matches), mean_goals = mean(goals), 
            sd_goals = sd(goals), mean_assists = mean(assists), sd_assists = sd(assists))
```

The average matches played during season 2018-2019 by these top 50 players is 17.34 and the standard deviation is 3.51.
The average goals scored during season 2018-2019 by these top 50 players is 5.22 and the standard deviation is 4.51.
The average assists made during season 2018-2019 by these top 50 players is 3.04 and the standard deviation is 2.73.

```{r yellow_cards, echo=FALSE}
players_combined %>%
  count(yellow_cards)
```

We can see that until now, 38 among the 50 players have got yellow cards during season 2018-2019, and the maximum yellow card number is 5.

```{r red_card, echo=FALSE}
players_combined %>%
  filter(red_cards!=0) %>%
  select(name, red_cards)
```

In season 2018-2019, there are 3 players who have already got red card in a match and being expelled: Kylian Mbappé, Cristiano Ronaldo, and Marcus Rashford.

```{r market_value, echo=FALSE}
players_combined %>%
  summarise(max_value = max(market_value), min_value = min(market_value))
players_combined %>%
  filter(market_value>=100) %>%
  summarise(value_above_100m = n())
```

In the dataset, the maximum player value is 180 million euro and the minimum player value is 65 million euro. There are 13 players in the dataset with a market value higher than 100 million euro.

### Visualizations

```{r position_goals, fig.width = 10, echo=FALSE}
players_combined %>%
  mutate(position = fct_reorder(position, goals)) %>% # The plot still does not reorder correctly
  ggplot(mapping = aes(x = position, y = goals)) +
  geom_col() +
  coord_flip() +
  labs(title = "Position and goals scored", 
       x = "Player Position", y = "Goals Scored")
```



```{r position_value, fig.width = 7, echo=FALSE}
players_combined %>%
  group_by(position) %>%
  summarise(mean_value = mean(market_value)) %>%
  mutate(position = fct_reorder(position, mean_value)) %>% # However, the same code here produces a correct reordering
  ggplot(mapping = aes(x = position, y = mean_value)) +
  geom_col() +
  coord_flip() +
  labs(title = "Position and market value", 
       x = "Player Position", y = "Player Market Value")
```

## Expected Results

1. We can also use existing player's profile to test the validity of our model.
2. We can compare the our model of player's market value with historical data so that we can obtain the trend of transfer market.

## Section 3. Data

```{r glimpse_data, echo=FALSE}
glimpse(players_combined)
dim(players_combined)
```

