---
title: "Football Exploration"
author: "team-devils"
date: "9/19/2018"
output: github_document
---

```{r load_packages, include = FALSE}
library(tidyverse)
library(broom)
library(rvest)
```


## Section 1. Introduction

We wish to explore how performance, physical attributes, and general characteristics of soccer players impact their market value during a season. Through visualizations and modeling, we hope to explore which characteristics are more and less determinant of strong performance and market value for a player, and whether patterns exist within goals, assists, age ranges, positions, and more. We plan to first minimally use visualizations to discover patterns, and then modeling to develop our central questions of which characteristics of players impact performance and value. 

Our data is a collection of the top 50 valuable players in the 2018-2019 European club season. The data comes from a professional German soccer statistics website titled "Transfermarkt", which is a website dedicated to tracking player's market values and performances. Each observation is a player, and includes the variables name, position, number of matches, number of goals scored, number of own goals, number of assists, number of yellow cards, number of red cards, number of substitutions on, number of substitutions off, and market value. The data was obtained from transfermarkt.com through web scraping tools learned from the course, and the web_scrape.R file can be found in this R project. We currently have 12 columns and 50 rows scraped, but may later during our analysis decide to obtain a few more columns and several more rows, depending on where our analysis leads us.

Source of data: Transfermarkt.com is a leading media in reporting soccer transfer news and they have connections with all the major leagues and clubs across Europe, South America, and Asia. The player statistics are generated after each match and analyzed by professional scouts and soccer analysts.

## Section 2. Data analysis plan

## Variables

Currently, there are 12 variables in our dataframe: name (name of players)
position (player's position on field)
age
matches (total appearances in season 2018-2019)
goals (goals scored in season 2018-2019)
own_goals (own goals scored in season 2018-2019)
assists (total assists made in season 2018-2019)
yellow_cards (yellow cards got in season 2018-2019)
red_cards (red cards got in season 2018-2019)
substituted_on (times of being substituted onto the field in season 2018-2019)
substituted_off (times of being substituted off the field in season 2018-2019)
market_value (up-to-dated value of the player in transfer market)

Here we use 10 variables as predictors (position, age, matches, goals, own_goals, assists, yellow_cards, red_cards, substituted_on, substituted_off). These variables are used to predict the final outcome variable ---- market_value.

## Preliminary Summary data and Visualization

```{r load-data, echo=FALSE}
players_combined <- read.csv("players_combined.csv")
glimpse(players_combined)
```

```{r summary-statistics, echo=FALSE}
players_combined %>%
  count(position)
```

We can see that the most common position among the top 50 players is Centre-Forward. Most players are attackers or midfielders instead of defenders.

```{r age}
players_combined <- players_combined %>%
  mutate(age_range = case_when(
    age>30  ~ "above 30",
    26<=age & age<=30     ~ "26-30",
    21<=age & age<=25     ~ "21-25",
    age<=20  ~ "under 20"
  ))
players_combined %>%
  count(age_range)
```

We have tried to classify the players by age. Most players age between 26 and 30. There are 3 players above 30 years old and there's also one player who is younger than 20.

```{r matches, echo=FALSE}
players_combined %>%
  summarise(mean_matches = mean(matches))
players_combined %>%
  summarise(sd_matches = sd(matches))
```

The average matches played during season 2018-2019 by these top 50 players is 17.34 and the standard deviation is 3.51.

```{r goals, echo=FALSE}
players_combined %>%
  summarise(mean_goals = mean(goals))
players_combined %>%
  summarise(sd_goals = sd(goals))
```

The average goals scored during season 2018-2019 by these top 50 players is 5.22 and the standard deviation is 4.51.

```{r assists, echo=FALSE}
players_combined %>%
  summarise(mean_assists = mean(assists))
players_combined %>%
  summarise(sd_assists = sd(assists))
```

The average assists made during season 2018-2019 by these top 50 players is 3.04 and the standard deviation is 2.73.

```{r yellow_cards, echo=FALSE}
players_combined %>%
  count(yellow_cards)
```

We can see that until now, 38 among the 50 players have got yellow cards during season 2018-2019, and the maximum yellow card number is 5.

```{r red_card, echo=FALSE}
players_combined %>%
  filter(red_cards!=0)
```

In season 2018-2019, there are 3 players who have already got red card in a match and being expelled: Kylian Mbappé, Cristiano Ronaldo, and Marcus Rashford.

```{r market_value}
players_combined %>%
  summarise(max_value = max(market_value))
players_combined %>%
  summarise(min_value = min(market_value))
players_combined %>%
  filter(market_value>=100) %>%
  summarise(value_above_100m = n())
```

In the dataset, the maximum player value is 180 million euro and the minimum player value is 65 million euro. There are 13 players in the dataset with a market value higher than 100 million euro.

## Section 3. Data
