---
title: "Building the Perfect Soccer Player"
author: "team-devils"
date: "Dec 14, 2018"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(20181213)
```

```{r load_packages, include = FALSE}
library(tidyverse)
library(broom)
library(rvest)
library(infer)
library(modelr)
library(knitr)
library(shiny)
library(tools)
library(shinythemes)
```



## Introduction

While its popularity may be somewhat subdued in the United States, soccer is not only one of the most widely followed sports in the world, but also a deeply significant shared cultural experience. According to the Global World Research Index, 3.2 billion individuals across the globe watched at least one portion of the 2018 FIFA World Cup. In July of 2018, Portuguese forward Cristiano Ronaldo transferred from Spanish club Real Madrid to Italian club Juventus in a €100 million deal. What combination of factors led to this one soccer player’s multi-million euro evaluation? Our team aims to discover what makes the most valuable soccer player. Here, the market value of a player is defined as the monetary amount that a club would pay in order to sign the player from his original club (also known as transfer). We will use data of players in the 2018-2019 season of the professional European club soccer leagues (majorly Premier League, La Liga, Serie A, Bundesliga, and France Ligue 1). We selected the information of players in this league over others due to the fact that besides the World Cup, the professional European leagues are the highest watched and most widely analyzed leagues in the world. Through visualizations exploring patterns between markers of success in the sport of soccer (such as number of assists, number of goals, etc.), markers of failure (such as number of yellow cards, number of red cards, number of own goals, etc.), and market value. In order to answer our central question, we plan to use linear regression and modeling to assist us in creating the most valuable soccer player. Our data is a collection of the top 500 most valuable players in the 2018-2019 European club season (data collected on 11/27/2018). We recognize that by using the top 500 players by market value (which is in and of itself a biased measure), we have a sample which is biased and may be right skewed in terms of market value. We will not be conducting hypothesis testing due to the former and we will explore possibly analyzing our data after using a transformation to account for the latter. The data comes from a professional German soccer statistics website titled “Transfermarkt”, which is a website dedicated to tracking players’ market values and performances. Transfermarkt.com is a leading medium in reporting soccer transfer news and they have connections with all of the major leagues and clubs across Europe, South America, and Asia. The player statistics are generated after each match and analyzed by professional scouts, soccer analysts and data scientists. Each observation is a player, and includes the variables name, position, number of matches, number of goals scored, number of own goals, number of assists, number of yellow cards, number of red cards, number of substitutions on, number of substitutions off, and market value. To better fit the model, we cleaned up the data by merging age and position into four categories. The data was obtained from transfermarkt.com through web scraping tools learned from the course.


## Data Analysis

```{r load-data, include = FALSE}
players1 <- read_csv("../data/players1.csv")
```

To get a general sense of the market values of all players in the 2018-2019 season, let's first create a histogram to visualize their distribution: 

```{r histogram, fig.width = 10, fig.cap = "Figure 1"}
players1 %>%
  ggplot(mapping = aes(market_value)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "Distribution of Player Market Values in 2018-2019 Season",
    x = "Player Market Values in 2018-2019 Season", 
    y = "Count"
  ) +
  theme_minimal()
```

We also performed a log transformation on the market values of players for comparison: 

```{r log_histogram, fig.width = 10, fig.cap = "Figure 2"}
players1 %>%
  ggplot(mapping = aes(log(market_value))) +
  geom_histogram(binwidth = 0.1) +
  labs(
    title = "Distribution of the Logarithm of Player Market Values in 2018-2019 Season",
    x = "Player Market Values in 2018-2019 season", 
    y = "Count"
  ) +
  theme_minimal()
```

Since the data we have chosen concerns only the top 500 valuable players, the dataset is right skewed instead of normally distributed. After the log transformation, the dataset remains right-skewed and does not provide any additional insights for the readers. Therefore, we decided to adhere to the original model and discard the log transformation.

In the Figure 1, we can see that the distribution of the top 500 players' market values in the 2018-2019 season is right-skewed, and that the most commonly occuring market value is around $23 million. In the summary statistics below, we can see that the mean of the market values is higher than the most commonly occuring market values due to the right skewedness of the data: 

```{r summary-stats}
players1 %>%
  summarise(mean = mean(market_value), median = median(market_value), sd = sd(market_value), min = min(market_value), max = max(market_value))

```

The median, however, seems to be affected to a lesser extent by the high valued outliers. In the boxplot below, the blue colored line represents the median, and the red line represents the mean: 

```{r boxplot, fig.height = 5, fig.width = 10, fig.cap = "Figure 3"}
players1 %>%
  ggplot(mapping = aes(y = market_value)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Boxplot of Player Market Values in 2018-2019 Season",
    y = "Player Market Values in 2018-2019 season"
  ) +
  theme_minimal() +
  geom_hline(yintercept = 35.3, color = "red") +
  geom_hline(yintercept = 25, color = "blue") +
  theme(axis.text.y=element_blank()) 
```

The boxplot also revealed that there are many outliers in the data. The table below shows the names and market values of the outlying soccer players: 

```{r outliers}
players1 %>%
  select(name, market_value) %>%
  filter(market_value > 75) %>%
  kable()

```

These players have extremely high market values compared to the rest of the players. Which raises the question: do they contribute much more than the other players? Let a player's contribution be the sum of the player's goals and assists, the following boxplots compare the contributions of the outlying players to that of the "normal" players: 

```{r outliers-plot, fig.width = 10, fig.cap = "Figure 4"}
players1 <- players1 %>%
  mutate(outlier = ifelse(
    market_value >75, T, F
  ), 
  contribution = goals + assists
  )

players1 %>%
  ggplot(mapping = aes(x = outlier, y = contribution, color = outlier)) +
  scale_color_manual(values = c("blue", "red")) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Player Contributions in 2018-2019 Season",
    x = "Outlier (True/False)", 
    y = "Count"
  ) +
  guides(color = "none") +
  coord_flip() +
  theme_minimal() 
```

We can see from the comparison above that the outlying soccer players with very high market values do contribute significantly more, therefore their high market value is justified.

Now we would like to investigate the relationship between a player's position and the number of goals they score. In order to reduce confusion, we condensed our data of 13 player positions into 4: Defender, Forward, Goalkeeper, and Midfielder. We found the average goals made by each position and average assists made by each position: 

```{r position-goals, fig.height = 4, fig.width = 10, fig.cap = "Figure 5"}
position_goals <- players1 %>%
  group_by(position_new) %>%
  summarise(goals_avg = mean(goals), assists_avg = mean(assists))

position_goals

position_goals %>%
  mutate(position_new = fct_reorder(position_new, goals_avg)) %>%
  ggplot(mapping = aes(x = position_new, y = goals_avg)) +
  geom_col(fill = "blue") +
  labs(
    title = "Average Goals Scored in Each Position in the 2018-2019 Season",
    x = "Position", 
    y = "Average Goals Scored"
  ) +
  coord_flip() +
  theme_minimal()
```

```{r position-assists, fig.height = 4, fig.width = 10, fig.cap = "Figure 6"}
position_goals %>%
  mutate(position_new = fct_reorder(position_new, assists_avg)) %>%
  ggplot(mapping = aes(x = position_new, y = assists_avg)) +
  geom_col(fill = "purple") +
  labs(
    title = "Average Assists in Each Position in the 2018-2019 Season",
    x = "Position", 
    y = "Average Assists"
  ) +
  coord_flip() +
  theme_minimal()
``` 

From the visualizations above, we can see that forward players have the most goals and assists on average, followed by midfielder, defender and goalkeeper. This is expected because forward players are the closest to the goal of the opposing team, making them the most likely to score and assist goals. Midfielders are the second closest, followed by defenders and goalkeepers. Interestingly, the average assists for goalkeeper is greater than 0. Let's find out which goalkeeper assisted a goal: 

```{r goal-assist}
players1 %>%
  select(name, position_new, assists) %>%
  filter(position_new == "Goalkeeper", assists > 0)
```

It looks like Ederson is the only goalkeeper who assisted a goal in the 2018-2019 season. 


### What Makes a Valuable Soccer Player? 

Below is an interactive scatterplot where we can plot all numerical variables against each other. The opacity of the data points are reduced so that we can better visualize overlapping data. We used this app to get a rough idea of which numerical variables are correlated with one another. 

```{r shiny-app, echo = F}
shinyApp(

ui <- fluidPage(
  theme = shinytheme("cyborg"),
  titlePanel("Interactive Player Stats Visualization"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        inputId = "y",
        label = "Y-axis:",
        choices = c(
          "Position" = "position",
          "Age" = "age",
          "Matches" = "matches",
          "Goals" = "goals",
          "Own Goals" = "own_goals",
          "Assists" = "assists",
          "Yellow Cards" = "yellow_cards",
          "Red Cards" = "red_cards",
          "Substituted On" = "substituted_on",
          "Substituted Off" = "substituted_off",
          "Market Value" = "market_value",
          "Age Range" = "age_range"
        ),
        selected = "market_value"
      ),
      selectInput(
        inputId = "x",
        label = "X-axis:",
        choices = c(
          "Position" = "position",
          "Age" = "age",
          "Matches" = "matches",
          "Goals" = "goals",
          "Own Goals" = "own_goals",
          "Assists" = "assists",
          "Yellow Cards" = "yellow_cards",
          "Red Cards" = "red_cards",
          "Substituted On" = "substituted_on",
          "Substituted Off" = "substituted_off",
          "Market Value" = "market_value",
          "Age Range" = "age_range"
        ),
        selected = "goals"
      ),
      selectInput(
        inputId = "z",
        label = "Color by:",
        choices = c(
          "Position" = "position",
          "Age" = "age",
          "Matches" = "matches",
          "Goals" = "goals",
          "Own Goals" = "own_goals",
          "Assists" = "assists",
          "Yellow Cards" = "yellow_cards",
          "Red Cards" = "red_cards",
          "Substituted On" = "substituted_on",
          "Substituted Off" = "substituted_off",
          "Market Value" = "market_value",
          "Age Range" = "age_range"
        ),
        selected = "age"
      )
    ),
    mainPanel(plotOutput(outputId = "scatterplot"))
  )
),

server <- function(input, output) {

  # Create scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = players1, aes_string(
      x = input$x, y = input$y,
      color = input$z
    )) +
      geom_point(alpha = 0.2) +
      labs(
        x = toTitleCase(str_replace_all(input$x, "_", " ")),
        y = toTitleCase(str_replace_all(input$y, "_", " ")),
        color = toTitleCase(str_replace_all(input$z, "_", " "))
      ) +
      scale_color_gradient(low="blue", high="red") +
      theme_minimal()
  })
},

options = list(height = 500)
)
```

To answer the question "What makes the a valuable soccer player?", we made a multiple linear regression based on all the variables we have in the players dataset. In soccer, a player's position is highly correlated with the number of goals and assists (e.g. forward positions score the most goals and make the most assists in general, while goalkeepers can seldom score a goal or make an assist). This is also evident from figures 5 and 6. From the Shiny app, we also saw that goals and assists are strongly correlated with a player's market value. Therefore, we decided to introduce two interactions between position/goals and position/assists into our multiple linear model. 

```{r predict}
linear_prediction  <- lm(market_value ~ -1 + factor(position_new) + age + matches + goals + own_goals +
                  assists + yellow_cards + red_cards + substituted_on +
                  substituted_off + factor(age_range) + factor(position_new) * goals + factor(position_new) * assists, data =   players1)
tidy(linear_prediction)

```

Next, we performed a backwards model selection based on AIC to identify the variables that significantly affect the players' market values. 

```{r backward-selection, include = FALSE}
selected_model <- step(linear_prediction, direction = "backward")
```

```{r selected-model}
tidy(selected_model)
```

After the model selection based on AIC, we can see that the new model has reduced the significant variables down to 15; variables "age", "own_goals", "yellow_cards", "red_cards", "substituted_on", and "substituted_off" have been eliminated.

```{r glance-AIC}
tibble(original_model = glance(linear_prediction)$AIC, 
       new_model = glance(selected_model)$AIC
) %>%
  kable()
```

The AIC value is lower compared to the full model (`r glance(linear_prediction)$AIC` to `r glance(selected_model)$AIC`), which means that the new model contains the variables that are the most likely to affect the market values. Taking a look at the adjusted R-squared values before and after backwards selection, our new model has a slightly higher adjusted R-squared value(`r glance(linear_prediction)$adj.r.squared` to `r glance(selected_model)$adj.r.squared`): 

```{r r-squared}
tibble(
  adj_r_squared_before = glance(linear_prediction)$adj.r.squared,
  adj_r_squared_after = glance(selected_model)$adj.r.squared
) %>%
  kable()
```

Also, in order to validate our model, we decide to take a look at the RMSE in a 10-fold test.

```{r cross_validation, include = FALSE}
test_cv <- crossv_kfold(players1, 10)
models <- map(test_cv$train, ~ selected_model)
rmses <- map2_dbl(models, test_cv$test, rmse)

```

```{r rmses}
rmses
```

In conclustion, according to the selected model, we can build a formula representing the market value prediction:

market_value = 15.9 - 0.607 * position_newForward + 22.7 * position_newGoalkeeper + 19.6 * position_newMidfielder + 0.406 * matches + 2.02 * goals + 0.885 * assists + 3.66 * age_range21-25 + 9.07 * age_range26-30 - 0.922 age_range30 and above + 0.937 * position_newForward:goals - 1.45 * position_newMidfielder:goals + 2.74 * position_newForward:assists + 23.4 * position_newGoalkeeper:assists - 0.366 * position_newMidfielder:assists

Now let's take a closer look at the model if we plug in different positions:

For Defenders :
market_value = 15.9 + 0.406 * matches + 2.02 * goals + 0.885 * assists + 3.66 * age_range21-25 + 9.07 * age_range26-30 - 0.922 age_range30 and above
Defenders' market values will increase if making more appearances and getting more goals and assists (though assists and goals are relatively unusual stuff for defenders). Also, 26-30 is the golden age range for a defender.

For Goalkeepers : 
market_value = 38.6 + 0.406 * matches + 2.02 * goals + 24.285 * assists + 3.66 * age_range21-25 + 9.07 * age_range26-30 - 0.922 age_range30 and above
Goalkeepers' market values will increase if making more appearances, aging between 26 and 30, and getting more goals and assists (Goalies! Try to run into opponent's box to make an assist even if it's NOT your zone).

For Midfielders : 
market_value = 35.5 + 0.406 * matches + 0.57 * goals + 0.519 * assists + 3.66 * age_range21-25 + 9.07 * age_range26-30 - 0.922 age_range30 and above
Midfielders' market values will increase if making more appearances, aging between 26 and 30, and getting more goals and assists. The coefficients of goals and assists for midfielders are lower than those of defenders because being high in these statistics actually compensate for that.

For Forwards : 
market_value = 15.293 + 0.406 * matches + 2.957 * goals + 3.625 * assists + 3.66 * age_range21-25 + 9.07 * age_range26-30 - 0.922 age_range30 and above
Forwards' market values will increase if making more appearances and getting more goals and assists. The starting intercept of forwards is negative. Still, the golden age of a forward is between 26 and 30. However, the Mr. Most Expensive, Kylian Mbappe (Forward), is only 19. What a bright young man!



## Conclusion

Your project goes here! Before you submit, make sure your chunks are turned off with `echo = FALSE`. 

You can add sections as you see fit. Make sure you have a section called Introduction at the beginning and a section called Conclusion at the end. The rest is up to you!
